Supporting Information: Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes.

### 2020-06-03 Ben_Manuscripts/hdphmm/Draft.tex
revised comments.
Addressed final comments

  In this work, we are particularly interested in the design of lyotropic liquid 
  crystal (LLC) monomers, a class of amphiphilic molecules whose ordered self-assembled
  phases can be cross-linked into mechanically strong membranes capable of highly 
  selective separations. Inverted hexagonal phase, or H\textsubscript{II}, LLC membranes
  are characterized by hexagonally packed, uniform-sized and straight pores, an ideal
  geometry for high throughput transport. Their pores are lined with the LLC monomer's
  functional groups which can potentially be designed to interact with solutes in a 
  chemically-specific manner. It would be extremely useful to experimentalists if we 
  could connect LLC monomer design with measurable quantities. This goal is complicated
  by the non-Brownian hopping and trapping behavior induced by the membrane's diverse
  topology and inhomogeneous structure resulting in subdiffusive behavior out to 
  multiple microseconds of simulation.~\cite{coscia_understanding_2019,coscia_chemically_2019}
  %MRS: citation missing for coscia_chemically_2019
  We have already advanced our understanding of LLC monomer design in our previous work
  by building stochastic time series models based on the chemical intuition gained
  from qualitative MD studies of solute transport in an H\textsubscript{II} phase LLC
  membrane.~\cite{coscia_chemically_2019,coscia_capturing_2020} We identified three 
  classes of trapping behavior which result in subdiffusion out to the microsecond 
  time scale. In our first approach to stochastically model this behavior, we treated 
  the system's dynamics as a sequence of anti-correlated hops between power law distributed periods of 
  entrapment, a framework called subordinated fractional Brownian or L\'evy
  motion.~\cite{thiel_weak_2014,teuerle_modeling_2013} In a second approach, we treated 
  solute motion as a Markov state model~\cite{pande_everything_2010} with state-dependent
  dynamics, where we parameterized the state transition probabilities between each of
  eight discrete states, defined by the observed trapping mechanisms, as well as the 
  solute dynamics within each of these states. We showed how one could use realizations
  of any stochastic model like these in order to predict macroscopic flux and selectivity.

@@

  \item Second, we want to know if the HDP-AR-HMM algorithm can help
  efficiently uncover underlying transport mechanisms which give rise to different 
  solute dynamical behavior. We cluster 
  %MRS:
  %similar state parameters 
  states with similar parameters
  across all
  trajectories in order to reduce the total state space and to 
  %MRS:
  identify
  %understand 
  which
  segments of the solute trajectories exhibit similar behavior. We use the 
  parameters of the clustered states identified by the HDP-AR-HMM in order to 
  infer dominant solute-membrane interactions and transport mechanisms. We 
  support our mechanistic hypotheses via quantitative comparison to various 
  physical membrane properties and solute-membrane interactions.

@@

  We studied transport of solutes in the H\textsubscript{II} phase using an
  atomistic molecular model of four pores in a monoclinic unit cell with 
  10\% water by weight. Approximately one third of the water molecules 
  occupy the tail region with the rest are near the pore center.~\cite{coscia_chemically_2019}
  %MRS: fix citation above.

@@

  We use the sticky HDP-AR-HMM in order to assign the time-correlated positional 
  fluctuations of solute center-of-mass MD trajectories to discrete dynamical modes while 
  simultaneously inferring the VAR parameters of those modes. We use the sticky model
  in order to encourage higher self-transition probabilities since we physically expect
  %MRS: maybe ``physically observe?'' instead of physically expect.
  solutes to stay trapped in states for appreciable amounts of time.

@@

  We clustered states with like parameter sets in order to reduce the state space to
  a more easily interpretable size 
  % CPC: perhaps keep hammering home not only reducing, but also reemphasize why absolute
  % value of some parameters really not of interest 
  %MRS: explain why parameters of single states don't give insight; more context needed?  Is CPC talking about
  %MRS: some some of the parameters not being interesting in general, not just the values for some states?
  and because the parameters of a single state alone does not always have enough context
  to provide meaningful physical insight. For each solute studied, we identified 
  approximately 300-500 independent states, each with separate VAR(1) parameters.
  Many of these states exhibit very similar dynamical behavior except their mean 
  levels in each dimension are different.

@@

  We clustered based on the eigenvalues of $A$ and $\Sigma$, the distance from the 
  pore center and the self-transition probabilities, $P_{ii}$, of each state. $A$ 
  and $\Sigma$ quantify the temporal correlation and size of fluctuations in a 
  given state. $P_{ii}$ is directly related to the expected time spent in each state.
  The distance from the closest pore center is expected to be physically meaningful
  since the membrane is radially inhomogeneous from each pore, transitioning from 
  hydrophilic pores to hydrophobic tails. However, the channels are isotropic along
  the pore axis so we do not cluster on the means in the $z$ dimension. Before carrying
  out the agglomerative clustering algorithm, we normalized each clustering dimension
  by its variance and shifted its mean to zero.

@@

  To cluster on $P_{ii}$, we first cast them in terms of the expected dwell times. The expected
  value of the number of sequential self-transitions is simply:
  \begin{equation}  
   E[dwell~time] = \frac{1}{1 - P_{ii}}
   \label{eqn:dwell_times}
  \end{equation}  
  This relationship implies that dwell times approach infinity as $P_{ii}$ approaches
  1. Therefore we found the most success with agglomerative clustering by linearizing this
  relationship and clustering on $-\log(1 - P_{ii})$.

@@

  We obtained estimates of clustered $A$ and $\Sigma$ parameters as well as a
  %MRS: check again agreement on $T$ and $P$ - you switched to $P$, right? Or are they 2 different things?
  consensus $T$ matrices, describing transitions between all possible clustered states 
  %MRS: do you explicitly say how to get the consensus transition matrices?
  exhibited by the ensemble of trajectories, using the inference component of 
  the HDP-AR-HMM algorithm. We applied the HDP-AR-HMM simultaneously to all 24
  zeroed solute trajectories with state sequences fixed according to the remapped 
  clustered state assignments. Simultaneous application of the HDP-AR-HMM to 
  all trajectories as we have done is only possible because the state sequence 
  is fixed. In cases where one does not know the state sequence and would like to 
  incorporate data from multiple trajectories, one can apply the beta process (BP)
  AR-HMM.~\cite{fox_sharing_2009} However the data requirements may be quite high
  for a good parameterization, which is why we elected to follow the clustering 
  approach described thus far.
  %BJC: I actually found 3 different ways do this. All giving the same results.
  % Method 1: described above - this is the most efficient way
  % Method 2: Link all 24 trajectories by adding a coffin state (basically a 
  % state that is a vector of zeros) between each traj
  % Method 3: Parameterize each trajectory independently, average together. T is
  % estimated by making a count matrix based on inputs from each trajectory

@@

  We reported the mean and $1 \sigma$ confidence intervals of the MSDs based on the 
  results of 200 bootstrap trials. Our bootstrapping procedure varied dependent on the
  method of trajectory generation:
  %MRS: this is a bit confusing.  Is there an expectation these should give consistent parameters?
  %MRS: the second is boostrapping an MSD over 1000 trajectories, and the first is boostrapping an MSD over 24 trajectories.
  %MRS: these should have different amounts of noise, making them hard to compare? Is this an issue?
  \begin{enumerate}[label={Method \theenumi :}, leftmargin=3.5\parindent]
  	\item For each bootstrap trial, we generated 24 stochastic trajectories, one from
  	each of the 24 independent HDP-AR-HMM models. The result of the trial is the mean
  	of the MSDs calculated from each trajectory constituting the ensemble. 
    \item We generated an ensemble of 1000 independent stochastic trajectories using
    the clustered parameter sets. For each trial, we calculated the MSD of 1000 
    trajectories sampled, with replacement, from the ensemble.
  \end{enumerate}

@@

  % BJC: Note that this figure is not finalized. It just takes a while to make 
  % because of bootstrapping and I keep forgetting to let it run overnight. I 
  % actually did it once, but had set save=False in the function I used in a jupyter notebook
  % In its final form, the errorbars will get smaller, and the means don't change by much.
  \begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{selectivity.pdf}
  \caption{Of the solutes studied, 
  %MRS: add qualifier.
  the models predict that 
  the current generation of H\textsubscript{II}
  LLC membranes may have the most success selectively removing acetic acid and,
  to a lesser extent, urea from solution containing either ethylene glycol or 
  methanol.
  }\label{fig:selectivity}
  \end{figure}

@@

  The distribution of types of hydrogen bonds and associated hydrogen bond lifetimes
  varies by predominant state, as seen in more detail in Figure~\ref{fig:hydrogen_bonding}.
  %MRS: what do you think about this? It maks it easier to read.
  \begin{itemize}
  \item States 1 and 2 primarily have hydrogen bonds to the 
  ester groups at the ends of the monomer tails and, to a smaller extent, the ether 
  linkages between the monomer head group and tails. 
  \item State 1 methanol molecules that
  donate hydrogen bonds to the tail oxygen atoms exhibit by far the longest hydrogen
  bond lifetimes among all states. 
  \item Methanol molecules in states 3 and 4 hydrogen 
  bond to all regions of the monomer relatively evenly, which is consistent with our earlier observation that
  states 3 and 4 are representative of behavior seen radially throughout the membrane
  pores. 
  \item State 3 has the lowest hydrogen bond and sodium association lifetimes which
  is consistent with its short dwell times. These are fleeting interactions since it
  is likely a high variance transitory state. 
  \item Finally, methanol molecules in state 5
  hydrogen bond almost exclusively to the carboxylate head groups and ether linkages 
  which is consistent with the low (close to the pore center) radial mean of the 
  cluster.
  \end{itemize}

  \begin{figure}
  \centering
  \begin{subfigure}{0.475\textwidth}
  \centering
  \includegraphics[width=0.6\textwidth, angle=0]{monomer_oxygens.png}
  \caption{}\label{fig:monomer_oxygens}
  \end{subfigure}
  \begin{subfigure}{0.475\textwidth}
  \includegraphics[width=\textwidth]{hbond_fractions.pdf}
  \caption{}\label{fig:hbond_fractions}
  \end{subfigure}
  \begin{subfigure}{0.475\textwidth}
  \includegraphics[width=\textwidth]{hbond_lifetimes.pdf}
  \caption{}\label{fig:hbond_lifetimes}
  \end{subfigure}
  \begin{subfigure}{0.475\textwidth}
  \includegraphics[width=\textwidth]{association_fraction_lifetimes.pdf}
  \caption{}\label{fig:association_fraction_lifetimes}
  \end{subfigure}
  \caption{(a) We analyzed hydrogen bonds donated by methanol to the highlighted oxygen
  atoms in three regions of the monomer: the carboxylate groups (blue), the ether linkages
  (green) and the tails (red). (b) For each predominant state (states 1--5 described in 
  Figure~\ref{fig:prevalence}), we plotted the fraction of the total time in that state 
  spent hydrogen bonding to various regions of the monomer, color coded to match (a).
  (c) For each type of hydrogen bond in each predominant state, we calculated the 95th percentile
  of hydrogen bond lifetimes. (d) We also analyzed the frequency and lifetimes of solute 
  association with sodium ions. Methanol associates with sodium less than 10\% of the total 
  time it occupies any of the predominant states.
  }
  \label{fig:hydrogen_bonding}
  \end{figure}

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{hbonds_assoc_summary.pdf}
  \caption{}\label{fig:hbonds_assoc_summary}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{rdf_summary.pdf}
  \caption{}\label{fig:rdf_summary}
  \end{subfigure}
  %BJC: Is it worth putting a bars with the solute MSDs on this plot
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{dwell_time_summary.pdf}
  \caption{}\label{fig:dwell_time_summary}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{density_comparison.pdf}
  \caption{}\label{fig:density_comparison}
  \end{subfigure}
  \caption{(a) Solutes hydrogen bond and associate with sodium ions
  to different degrees. Sometimes, they participate in both interactions at
  once. (b) The radial means of the unclustered states weighted
  by the length of time spent in that state. Most solutes spend the majority of
  their time close to the pore center. Only methanol has a significant density
  far from the pore center. (c) The 95\textsuperscript{th} percentile of dwell
  times, collected by measuring the length of each disjoint segment according 
  to the clustered state sequences, is inversely related to the solute MSDs.
  %MRS: this can be explained in the text.
  %after a 1000 ns time lag.
  (d) The expected dwell time of solutes increases in states with high local 
  number density of surrounding heavy atoms.
  }\label{fig:summaries}
  \end{figure}

@@

  \begin{figure}[!htb]
  \centering
  \includegraphics[width=3.25in]{hdphmm_TOC.pdf}
  \end{figure}

>>>>

  In this work, we are particularly interested in the design of lyotropic liquid 
  crystal (LLC) monomers, a class of amphiphilic molecules whose ordered self-assembled
  phases can be cross-linked into mechanically strong membranes capable of highly 
  selective separations. Inverted hexagonal phase, or H\textsubscript{II}, LLC membranes
  are characterized by hexagonally packed, uniform-sized and straight pores, an ideal
  geometry for high throughput transport. Their pores are lined with the LLC monomer's
  functional groups which can potentially be designed to interact with solutes in a 
  chemically-specific manner. It would be extremely useful to experimentalists if we 
  could connect LLC monomer design with measurable quantities. This goal is complicated
  by the non-Brownian hopping and trapping behavior induced by the membrane's diverse
  topology and inhomogeneous structure resulting in subdiffusive behavior out to 
  multiple microseconds of simulation.~\cite{coscia_understanding_2019,coscia_chemically_2019}

  We have already advanced our understanding of LLC monomer design in our previous work
  by building stochastic time series models based on the chemical intuition gained
  from qualitative MD studies of solute transport in an H\textsubscript{II} phase LLC
  membrane.~\cite{coscia_chemically_2019,coscia_capturing_2020} We identified three 
  classes of trapping behavior which result in subdiffusion out to the microsecond 
  time scale. In our first approach to stochastically model this behavior, we treated 
  the system's dynamics as a sequence of anti-correlated hops between power law distributed periods of 
  entrapment, a framework called subordinated fractional Brownian or L\'evy
  motion.~\cite{thiel_weak_2014,teuerle_modeling_2013} In a second approach, we treated 
  solute motion as a Markov state model~\cite{pande_everything_2010} with state-dependent
  dynamics, where we parameterized the state transition probabilities between each of
  eight discrete states, defined by the observed trapping mechanisms, as well as the 
  solute dynamics within each of these states. We showed how one could use realizations
  of any stochastic model like these in order to predict macroscopic flux and selectivity.

@@

  \item Second, we want to know if the HDP-AR-HMM algorithm can help
  efficiently uncover underlying transport mechanisms which give rise to different 
  solute dynamical behavior. We cluster states with similar parameters
  across all trajectories in order to reduce the total state space and to 
  identify which segments of the solute trajectories exhibit similar behavior. We use the 
  parameters of the clustered states identified by the HDP-AR-HMM in order to 
  infer dominant solute-membrane interactions and transport mechanisms. We 
  support our mechanistic hypotheses via quantitative comparison to various 
  physical membrane properties and solute-membrane interactions.

@@

  We studied transport of solutes in the H\textsubscript{II} phase using an
  atomistic molecular model of four pores in a monoclinic unit cell with 
  10\% water by weight. Approximately one third of the water molecules 
  occupy the tail region with the rest are near the pore center.~\cite{coscia_chemically_2019}

@@

  We use the sticky HDP-AR-HMM in order to assign the time-correlated positional 
  fluctuations of solute center-of-mass MD trajectories to discrete dynamical modes while 
  simultaneously inferring the VAR parameters of those modes. We use the sticky model
  in order to encourage higher self-transition probabilities since we physically observe
  solute entrapment in states for appreciable amounts of time.

@@

  We clustered states with like parameter sets in order to reduce the state space to
  a more easily interpretable size and because the parameters of a single state 
  alone does not always have enough context to provide meaningful physical insight.
  For each solute studied, we identified approximately 300-500 independent states, 
  each with separate VAR(1) parameters. Many of these states exhibit very similar 
  dynamical behavior except their mean levels in each dimension are different.

@@

  We clustered based on the eigenvalues of $A$ and $\Sigma$, the distance from the 
  pore center and the self-transition probabilities, $T_{ii}$, of each state. $A$ 
  and $\Sigma$ quantify the temporal correlation and size of fluctuations in a 
  given state. $T_{ii}$ is directly related to the expected time spent in each state.
  The distance from the closest pore center is expected to be physically meaningful
  since the membrane is radially inhomogeneous from each pore, transitioning from 
  hydrophilic pores to hydrophobic tails. However, the channels are isotropic along
  the pore axis so we do not cluster on the means in the $z$ dimension. Before carrying
  out the agglomerative clustering algorithm, we normalized each clustering dimension
  by its variance and shifted its mean to zero.

@@

  To cluster on $T_{ii}$, we first cast them in terms of the expected dwell times. The expected
  value of the number of sequential self-transitions is simply:
  \begin{equation}  
   E[dwell~time] = \frac{1}{1 - T_{ii}}
   \label{eqn:dwell_times}
  \end{equation}  
  This relationship implies that dwell times approach infinity as $T_{ii}$ approaches
  1. Therefore we found the most success with agglomerative clustering by linearizing this
  relationship and clustering on $-\log(1 - T_{ii})$.

@@

  We obtained estimates of clustered $A$ and $\Sigma$ parameters as well as
  consensus $T$ matrices, describing transitions between all possible clustered states 
  exhibited by the ensemble of trajectories, using the inference component of 
  the HDP-AR-HMM algorithm. We applied the HDP-AR-HMM simultaneously to all 24
  zeroed solute trajectories with state sequences fixed according to the remapped 
  clustered state assignments. This allowed us to parameterize states shared 
  across multiple trajectories and to easily generate a single transition matrix. 
  Simultaneous application of the HDP-AR-HMM to all trajectories as we have done 
  is only possible because the state sequence is fixed. In cases where one does not 
  know the state sequence and would like to infer by incorporating data from 
  multiple trajectories, one can apply the beta process (BP) AR-HMM.~\cite{fox_sharing_2009} 
  However the data requirements may be quite high for a good parameterization, 
  which is why we elected to follow the clustering approach described thus far.
  %BJC: I actually found 3 different ways do this. All giving the same results.
  % Method 1: described above - this is the most efficient way
  % Method 2: Link all 24 trajectories by adding a coffin state (basically a 
  % state that is a vector of zeros) between each traj
  % Method 3: Parameterize each trajectory independently, average together. T is
  % estimated by making a count matrix based on inputs from each trajectory

@@

  We reported the mean and $1 \sigma$ confidence intervals of the MSDs based on the 
  results of 200 bootstrap trials. Our bootstrapping procedure varied dependent on the
  method of trajectory generation:
  %MRS: this is a bit confusing.  Is there an expectation these should give consistent parameters?
  %MRS: the second is boostrapping an MSD over 1000 trajectories, and the first is boostrapping an MSD over 24 trajectories.
  %MRS: these should have different amounts of noise, making them hard to compare? Is this an issue?
  %BJC: added some clarification to Method 2
  \begin{enumerate}[label={Method \theenumi :}, leftmargin=3.5\parindent]
  	\item For each bootstrap trial, we generated 24 stochastic trajectories since we
  	are constrained to sample with replacement from the 24 independent HDP-AR-HMM models.
    The result of each trial is the mean of the MSDs of the trajectories
    constituting the bootstrapped ensemble. 
    \item Since trajectories generated by Method 2 are based on the clustered parameter 
    set used to define a single model, we can bootstrap on trajectory ensembles of
    arbitrary size. For each bootstrap trial, we generated an ensemble of 1000 independent
    stochastic trajectories, calculated the MSD of each and then sampled, with replacement,
    from the ensemble. This larger ensemble allows us to achieve lower uncertainties with
    the same number of bootstrap trials as Method 1. 
  \end{enumerate}

@@

  % BJC: Note that this figure is not finalized. It just takes a while to make 
  % because of bootstrapping and I keep forgetting to let it run overnight. I 
  % actually did it once, but had set save=False in the function I used in a jupyter notebook
  % In its final form, the errorbars will get smaller, and the means don't change by much.
  \begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{selectivity2.pdf}
  \caption{Of the solutes studied, the models predict that the current generation 
  of H\textsubscript{II} LLC membranes may have the most success selectively removing 
  acetic acid and, to a lesser extent, urea from solution containing either ethylene
  glycol or methanol.
  }\label{fig:selectivity}
  \end{figure}

@@

  The distribution of types of hydrogen bonds and associated hydrogen bond lifetimes
  varies by predominant state, as seen in more detail in Figure~\ref{fig:hydrogen_bonding}.
  \begin{itemize}
	\item States 1 and 2 primarily have hydrogen bonds to the ester groups at the ends
	of the monomer tails and, to a smaller extent, the ether linkages between the 
	monomer head group and tails. 
    \item State 1 methanol molecules that donate hydrogen bonds to the tail oxygen 
    atoms exhibit by far the longest hydrogen bond lifetimes among all states. 
    \item Methanol molecules in states 3 and 4 hydrogen bond to all regions of the
    monomer relatively evenly, which is consistent with our earlier observation that
    states 3 and 4 are representative of behavior seen radially throughout the membrane
    pores. 
    \item State 3 has the lowest hydrogen bond and sodium association lifetimes which
    is consistent with its short dwell times. These are fleeting interactions since it
    is likely a high variance transitory state. 
    \item Finally, methanol molecules in state 5 hydrogen bond almost exclusively 
    to the carboxylate head groups and ether linkages which is consistent with the 
    low (close to the pore center) radial mean of the cluster.
  \end{itemize}

  \begin{figure}
  \centering
  \begin{subfigure}{0.475\textwidth}
  \centering
  \includegraphics[width=0.6\textwidth, angle=0]{monomer_oxygens.pdf}
  \caption{}\label{fig:monomer_oxygens}
  \end{subfigure}
  \begin{subfigure}{0.475\textwidth}
  \includegraphics[width=\textwidth]{hbond_fractions.pdf}
  \caption{}\label{fig:hbond_fractions}
  \end{subfigure}
  \begin{subfigure}{0.475\textwidth}
  \includegraphics[width=\textwidth]{hbond_lifetimes.pdf}
  \caption{}\label{fig:hbond_lifetimes}
  \end{subfigure}
  \begin{subfigure}{0.475\textwidth}
  \includegraphics[width=\textwidth]{association_fraction_lifetimes.pdf}
  \caption{}\label{fig:association_fraction_lifetimes}
  \end{subfigure}
  %y-axis label being cut off?
  \caption{(a) We analyzed hydrogen bonds donated by methanol to the highlighted oxygen
  atoms in three regions of the monomer: the carboxylate groups (blue), the ether linkages
  (green) and the tails (red). (b) For each predominant state (states 1--5 described in 
  Figure~\ref{fig:prevalence}), we plotted the fraction of the total time in that state 
  spent hydrogen bonding to various regions of the monomer, color coded to match (a).
  (c) For each type of hydrogen bond in each predominant state, we calculated the 95th percentile
  of hydrogen bond lifetimes. (d) We also analyzed the frequency and lifetimes of solute 
  association with sodium ions. Methanol associates with sodium less than 10\% of the total 
  time it occupies any of the predominant states.
  }
  \label{fig:hydrogen_bonding}
  \end{figure}

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{hbonds_assoc_summary.pdf}
  \caption{}\label{fig:hbonds_assoc_summary}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{rdf_summary.pdf}
  \caption{}\label{fig:rdf_summary}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{dwell_time_summary.pdf}
  \caption{}\label{fig:dwell_time_summary}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{density_comparison.pdf}
  \caption{}\label{fig:density_comparison}
  \end{subfigure}
  \caption{(a) Solutes hydrogen bond and associate with sodium ions
  to different degrees. Sometimes, they participate in both interactions at
  once. (b) The radial means of the unclustered states weighted
  by the length of time spent in that state. Most solutes spend the majority of
  their time close to the pore center. Only methanol has a significant density
  far from the pore center. (c) The 95\textsuperscript{th} percentile of dwell
  times, collected by measuring the length of each disjoint segment according 
  to the clustered state sequences, is inversely related to the solute MSDs.
  (d) The expected dwell time of solutes increases in states with high local 
  number density of surrounding heavy atoms.
  }\label{fig:summaries}
  \end{figure}

@@

  \begin{figure}[!htb]
  \centering
  \includegraphics[width=3.25in]{hdphmm_TOC_v2.pdf}
  \end{figure}

### 2020-06-03 Ben_Manuscripts/hdphmm/Supporting_Information.tex
Addressed final comments

  At extremely short time lags, the MSD is generally over-predicted by the HDP-AR-HMM
  (see Figure~\ref{fig:unclustered_msd_shortlag}) because the VAR(1) model assumes
  multivariate Gaussian noise while the MD data suggests the noise is better 
  modeled by more general L\'evy stable noise. The first time lag of the MSD is
  equivalent to the variance of all observed fluctuations. 
  The HDP-AR-HMM simulation appears to over-estimate the variance because the actual 
  distribution of solute fluctuations has 
  %MRS: heavy tails should mean larger variance? Clarify?
  heavy tails with a smaller variance (Figure~\ref{fig:emission_widths}).
  The total distribution of hops is expected to be heavy tailed because it is a
  combination of hop distributions from many states with different variances. More
  interestingly, in Figure ~\ref{fig:state_emission_widths}, we show that the 
  emission distribution of a single state picked out by the HDP-AR-HMM is fit better by
  a more general L\'evy distribution while the emissions from HDP-AR-HMM realizations are
  Gaussian, as expected. An improvement to the HDP-AR-HMM that may allow dynamics more 
  faithful to MD would use an autoregressive model with L\'evy stable noise. However,
  this would be less computationally feasible due to lack of conjugate priors.

>>>>

  At extremely short time lags, the MSD is generally over-predicted by the HDP-AR-HMM
  (see Figure~\ref{fig:unclustered_msd_shortlag}) because the VAR(1) model assumes
  multivariate Gaussian noise while the MD data suggests the noise is better 
  modeled by more general L\'evy stable noise. The first time lag of the MSD is
  equivalent to the variance of all observed fluctuations. 
  The HDP-AR-HMM simulation appears to over-estimate the width of the hop distribution
  because the actual distribution of solute fluctuations has 
  heavy tails with a smaller width (Figure~\ref{fig:emission_widths}).
  The total distribution of hops is expected to be heavy tailed because it is a
  combination of hop distributions from many states with different variances. More
  interestingly, in Figure ~\ref{fig:state_emission_widths}, we show that the 
  emission distribution of a single state picked out by the HDP-AR-HMM is fit better by
  a more general L\'evy distribution while the emissions from HDP-AR-HMM realizations are
  Gaussian, as expected. An improvement to the HDP-AR-HMM that may allow dynamics more 
  faithful to MD would use an autoregressive model with L\'evy stable noise. However,
  this would be less computationally feasible due to lack of conjugate priors.
